Icicle: write once, run once

### 2016-06-08 2016fhpc/section/04-Benchmarks.tex
Work on related work section

At Ambiata we are currently using Icicle in production over medium-sized data sets that fit on a single disk. We are also currenly implementing a scheduler to distribute larger data sets across multiple nodes. The data we are working with is tens of terabytes compressed, which in 2016 does not fit on a single disk. However, each row has a natural primary key and the features we need to compute depend only on the data within single key groups, which makes the workload very easy to distribute.

@@

We ran all the Unix utilities with unicode decoding disabled using @LANG=C LC_COLLATE@ for maximum performance. The input data does not contain unicode characters. 
We used an Amazon EC2 @c3.8xlarge@ with 32 CPUs, 60GB of RAM, and striped, RAIDed SSD storage. Icicle significantly outperformed R, and the single-threaded version was on par with @wc@, while only a little slower than @grep@. This is despite the fact that the Icicle queries perform more computational work than than @wc@ and @grep@. By using multiple processors, we were able to scale up to perform as well as @cat@, approaching the disk speed.
The R code is single threaded and would require at least 150 processors to reach similar speeds, assuming perfect scaling.
These results give us confidence that our distributed implementation will be fast as well as scalable~\cite{mcsherry2015scalability}.

>>>>

At Ambiata we are currently using Icicle in production over medium-sized data sets that fit on a single disk. We are also currently implementing a scheduler to distribute larger data sets across multiple nodes. The data we are working with is tens of terabytes compressed, which in 2016 does not fit on a single disk. However, each row has a natural primary key and the features we need to compute depend only on the data within single key groups, which makes the workload very easy to distribute.

@@

We ran all the Unix utilities with unicode decoding disabled using @LANG=C LC_COLLATE@ for maximum performance. The input data does not contain unicode characters. 
We used an Amazon EC2 @c3.8xlarge@ with 32 CPUs, 60GB of RAM, and striped, RAIDed SSD storage. The fused Icicle version significantly outperformed the R version of the queries, and the single-threaded version was on par with @wc@, while only a little slower than @grep@. This is despite the fact that the Icicle queries perform more computational work than than @wc@ and @grep@. By using multiple processors, we were able to scale up to perform as well as @cat@, approaching the disk speed.
The R code is single threaded and would require at least 150 processors to reach similar speeds, assuming perfect scaling.
These results give us confidence that our distributed implementation will be fast as well as scalable~\cite{mcsherry2015scalability}.

### 2016-06-08 2016fhpc/section/XX-Related.tex
Work on related work section

%!TEX root = ../Main.tex
\eject
\section{Conclusion}
\label{s:Conclusion}
This paper has introduced Icicle, a streaming query language.
The streaming, single-pass nature of Icicle allows all queries over the same table to be fused into a single loop over the data.
Icicle's modal type system allows incremental computation, ensuring that queries give the same value regardless of how the input is sliced up, and when the increments are performed.
The modal types encode when computations are available, and disallows using the results of folds before they are finished.

In Icicle there is only one stream, the input table, which cannot be directly referred to or modified.
This approach is simpler but less expressive than existing synchronous data flow languages such as Lucy~\cite{mandel2010lucy}, as well as our prior work on flow fusion~\cite{lippmeier2013data}.
These systems use clock types and causal analysis to ensure programs can be executed with bounded buffers~\cite{stephens1997survey}.
Stream combinators with multiple inputs, such as @zip@, are assigned types that require their stream arguments to have the same clock --- meaning that elements always arrive in lockstep and the combinators themselves do not need to perform their own buffering.

@@

It is well known that streaming dataflow languages correspond to Kahn process networks, and synchronous languages to restricted networks with bounded buffers~\cite{johnston2004advances}.
Kahn process networks have also been touted as a more flexible and intuitive abstraction than MapReduce~\cite{vrba2009kahn}.
Synchronous streaming dataflow languages such as Icicle and Lucy, then, should prove a useful abstraction for distributed computing.

>>>>

%!TEX root = ../Main.tex
\eject
\section{Related Work}
\label{s:Conclusion}

% This paper has introduced Icicle, a streaming query language. The streaming, single-pass nature of Icicle allows all queries over the same table to be fused into a single loop over the data.
% Icicle's modal type system allows incremental computation, ensuring that queries give the same value regardless of how the input is sliced up, and when the increments are performed.
% The modal types encode when computations are available, and disallows using the results of folds before they are finished.
% \ben{the above paragraph doesn't add any new information}.

In Icicle there is only one stream, the input table, which is implicit in the body of the queries. This approach is intentionally simpler than existing synchronous data flow languages such as Lucy~\cite{mandel2010lucy}, as well as our prior work on flow fusion~\cite{lippmeier2013data}.
These latter systems use clock types and causal analysis to ensure that programs can be executed with bounded buffers~\cite{stephens1997survey}. Synchronous data flow languages implement Kahn networks that are restricted to use bounded buffering~\cite{johnston2004advances}.
In such languages, stream combinators with multiple inputs, such as @zip@, are assigned types that require their stream arguments to have the same clock --- meaning that elements always arrive in lockstep and the combinators themselves do not need to perform their own buffering. In Icicle the fact that the input stream is implicit and distributed to all combinators means that we can forgo clock analysis. All queries in a program execute in lock-step on the same element at the same moment, which ensures that fusion is a simple matter of concatenating the components of the loop anatomy of each query.

@@

% It is well known that streaming dataflow languages correspond to Kahn process networks.

% Kahn process networks have also been touted as a more flexible and intuitive abstraction than MapReduce~\cite{vrba2009kahn}.
% This doesn't say anything about Icicle.

% Synchronous streaming dataflow languages such as Icicle and Lucy, then, should prove a useful abstraction for distributed computing.
